DOpElib: A Goal Oriented Software Library for Computing PDEs and Optimization Problems

### 2012-11-07 Paper_May_2012/main.tex
Habe versucht in den Seiten 6-8 etwas zu entzerren. Ausserdem gefallen mir die Paragraphen-Ueberschriften nicht, da die kursiv und sehr klein sind. Am liebsten wuerde ich die subsections aufwerten (Section 2.1: Solving of PDEs, Section 2.2: Solving of nonstationary PDEs, Section 2.3: solving of OPT. Dann hatten wir die subsection-Ueberschriften fuer localpde, etc. frei was momentan als Paragraph versteckt ist

%IntegratorDataContainer<VECTOR, 2> idc(quad_formula,
%                                         face_quad_formula);
%The \texttt{IntegratorDataContainer} class just makes the constructor interface of 
Then, the two characteristic features \texttt{LocalPDE} and \texttt{LocalFunctional} objects come into play.
\begin{lstlisting}
LocalPDE<VECTOR> local_pde;
LocalFunctional<VECTOR> local_functional;
\end{lstlisting}
These hold the information of the PDE and the functional under consideration. Further explication is given below because
the definition of this classes is done spererately in \texttt{localpde.h} resp.   \texttt{localfunctional.h}.

In the next step, we define an object of type \texttt{SpaceTimeHandler}, which handles all the things connected with the triangulation as well as the finite element space. This class is also capable of organizing degrees of freedom in time-dependent problems, hence the name. 
\begin{lstlisting}
  SpaceTimeHandler<TRIANGULATION, FE> dof_handler(triangulation,
                                                  finite_element);
\end{lstlisting}
There exist different versions of this class for forward as well as optimization problems.

Now we have everything at hand to create an instantiation of type \texttt{PDEProblemContainer}. This basically represents the disrectized problem, and bundles things like boundary conditions, the PDE, the functional and the discretization. First, we give the constructor the \texttt{local\_pde} and the \texttt{dof\_handler}.
\begin{lstlisting}
  PDEProblemContainer<LocalPDE> prob_container(local_pde,
                                               dof_handler);
\end{lstlisting}
After that, we add the functional under consideration and, because it is a boundaryfunctional, we specify the part of the boundary on which the functional operates (this is done via a 'boundarycolor', $1$ represents $\Gamma_{circ}$).
\begin{lstlisting}
  prob_container.AddFunctional(&local_functional);
  prob_container.SetBoundaryFunctionalColors(1);
\end{lstlisting}
Next, we incorporate the dirichlet data, which is given by the class \texttt{DirichleData}. Then we have to tell the \texttt{prob\_container} for which components and boundaries we prescribe the dirichlet conditions. This is done via a boundary-color (0 stands for $\Gamma_D$) and a \texttt{std::vector<bool> comp\_mask = (true, true, false)}.
\begin{lstlisting}
  DirichletData dd;
  prob_container.SetDirichletBoundaryColors(0, comp_mask, dd);
\end{lstlisting}
Now we create the object which steers the whole solution process, the previosly mentioned \texttt{StatPDEProblem}. It takes care of solving the discretized equation and evaluating the given functionals. The class depends actually on some template-parameters, so for the sake of abbreviation, we defined a typedef at the beginning of the \texttt{main.cc}-file.
\begin{lstlisting}
typedef StatPDEProblem<NEWTONSOLVER,
           PDEProblemContainer<LocalPDE>, VECTOR, 2> SSOLVER;
\end{lstlisting}
Here, \texttt{NEWTONSOLVER} itself is an abbreviation of
\begin{lstlisting}
typedef NewtonSolver<LINEARSOLVER, VECTOR, 2> NEWTONSOLVER;
\end{lstlisting}
The first template parameter of \texttt{NewtonSolver} describes which linear solver we use in this example. We opted for a direct solver, so: 
\begin{lstlisting}
typedef DirectLinearSolverWithMatrix<MATRIX,
            VECTOR, 2> LINEARSOLVER;
\end{lstlisting}
With the type settled, we create the object \texttt{solver}, which needs the discretization (represented here by the \texttt{prob\_container}), the parameter reader as well as the quadrature rules.
\begin{lstlisting}
  SSOLVER solver(prob_container, pr,
                    quad_formula, face_quad_formula);
\end{lstlisting}
%Finally, the output handler is initialized via
%\begin{lstlisting}
%  P.RegisterOutputHandler(&out);
%  solver.RegisterOutputHandler(&out);
%\end{lstlisting}
Finally, the programm is started by
\begin{lstlisting}
  solver.ReInit();
  solver.ComputeReducedFunctionals();
\end{lstlisting}

@@

In addition to stationary problems, the implementation
of the time-derivative must be considered which is done
in 
\begin{lstlisting}
  void CellTimeEquation(...)
   {
     // implement (u^{n+1},\phi) 
   }
\end{lstlisting}
and it is done in a straight-forward way:
\begin{lstlisting}
  for (all quadrature points per cell q)
   {
     Tensor<1,2> u = ...;

@@

\subsection{Compliance minimization}
In this example, we consider the compliance minimization of a standard MBB-Beam. 
The thickness variable is allowed to take intermediate value apart from $0$ and $1$,
i.e., we consider a variable thickness sheet. The discretization is done 
using $Q_2$ elements for the displacement and discontinuous $P_0$ elements for the 
thickness.
\begin{figure}
\centering
\includegraphics[width=1.\textwidth, viewport=150 0 1024 500, clip]{Pictures/MBB.png}
\caption{Thickness distribution for the MBB-beam minimum compliance problem.} 
\label{res:mbb}
\end{figure}
A typical thickness distribution is shown in Figure~\ref{res:mbb}.

>>>>

%IntegratorDataContainer<VECTOR, 2> idc(quad_formula,
%                                         face_quad_formula);
%The \texttt{IntegratorDataContainer} class just makes the constructor interface of 
Then, the two characteristic features \texttt{LocalPDE} and \texttt{LocalFunctional} objects come into play:
\begin{lstlisting}

  LocalPDE<VECTOR> local_pde;
  LocalFunctional<VECTOR> local_functional;

\end{lstlisting}
These hold the information of the PDE and the functional under consideration. Further explication is given below because
the definition of this classes is done spererately in \texttt{localpde.h} resp.   \texttt{localfunctional.h}.

In the next step, we define an object of type \texttt{SpaceTimeHandler}, which handles all the things connected with the triangulation as well as the finite element space. This class is also capable of organizing degrees of freedom in time-dependent problems, hence the name. 
\begin{lstlisting}

  SpaceTimeHandler<TRIANGULATION, FE> dof_handler(triangulation,
                                                  finite_element);

\end{lstlisting}
There exist different versions of this class for forward as well as optimization problems.

Now we have everything at hand to create an instantiation of type \texttt{PDEProblemContainer}. This basically represents the disrectized problem, and bundles things like boundary conditions, the PDE, the functional and the discretization. First, we give the constructor the \texttt{local\_pde} and the \texttt{dof\_handler}.
\begin{lstlisting}

  PDEProblemContainer<LocalPDE> prob_container(local_pde,
                                               dof_handler);
\end{lstlisting}
After that, we add the functional under consideration and, because it is a boundaryfunctional, we specify the part of the boundary on which the functional operates (this is done via a 'boundarycolor', $1$ represents $\Gamma_{circ}$).
\begin{lstlisting}

  prob_container.AddFunctional(&local_functional);
  prob_container.SetBoundaryFunctionalColors(1);

\end{lstlisting}
Next, we incorporate the dirichlet data, which is given by the class \texttt{DirichleData}. Then we have to tell the \texttt{prob\_container} for which components and boundaries we prescribe the dirichlet conditions. This is done via a boundary-color (0 stands for $\Gamma_D$) and a \texttt{std::vector<bool> comp\_mask = (true, true, false)}.
\begin{lstlisting}

  DirichletData dd;
  prob_container.SetDirichletBoundaryColors(0, comp_mask, dd);

\end{lstlisting}
Now we create the object which steers the whole solution process, the previosly mentioned \texttt{StatPDEProblem}. It takes care of solving the discretized equation and evaluating the given functionals. The class depends actually on some template-parameters, so for the sake of abbreviation, we defined a typedef at the beginning of the \texttt{main.cc}-file.
\begin{lstlisting}

  typedef StatPDEProblem<NEWTONSOLVER,
           PDEProblemContainer<LocalPDE>, VECTOR, 2> SSOLVER;

\end{lstlisting}
Here, \texttt{NEWTONSOLVER} itself is an abbreviation of
\begin{lstlisting}

  typedef NewtonSolver<LINEARSOLVER, VECTOR, 2> NEWTONSOLVER;

\end{lstlisting}
The first template parameter of \texttt{NewtonSolver} describes which linear solver we use in this example. We opted for a direct solver, so: 
\begin{lstlisting}

  typedef DirectLinearSolverWithMatrix<MATRIX,
            VECTOR, 2> LINEARSOLVER;

\end{lstlisting}
With the type settled, we create the object \texttt{solver}, which needs the discretization (represented here by the \texttt{prob\_container}), the parameter reader as well as the quadrature rules.
\begin{lstlisting}

  SSOLVER solver(prob_container, pr,
                    quad_formula, face_quad_formula);

\end{lstlisting}
%Finally, the output handler is initialized via
%\begin{lstlisting}
%  P.RegisterOutputHandler(&out);
%  solver.RegisterOutputHandler(&out);
%\end{lstlisting}
Finally, the programm is started by
\begin{lstlisting}

  solver.ReInit();
  solver.ComputeReducedFunctionals();

\end{lstlisting}

@@

In addition to stationary problems, the implementation
of the time-derivative must be considered which is done
in 
\begin{lstlisting}
  void CellTimeEquation(...)
   {
     local_cell_vector(i) += 
         scale * (u^{n+1},\phi) + etc. 
   }
\end{lstlisting}
and it is done in a straight-forward way:
\begin{lstlisting}
  for (all quadrature points per cell q)
   {
     Tensor<1,2> u = ...;

@@

\newpage
\subsection{Compliance minimization}
In this example, we consider the compliance minimization of a standard MBB-Beam. 
The thickness variable is allowed to take intermediate value apart from $0$ and $1$,
i.e., we consider a variable thickness sheet. The discretization is done 
using $Q_2$ elements for the displacement and discontinuous $P_0$ elements for the 
thickness.
\begin{figure}
\centering
\includegraphics[width=1.\textwidth, viewport=150 0 1024 500, clip]{Pictures/MBB.png}
\caption{Thickness distribution for the MBB-beam minimum compliance problem.} 
\label{res:mbb}
\end{figure}
A typical thickness distribution is shown in Figure~\ref{res:mbb}.
